---
title: "Proximal/Distal Analysis"
author: "Gustafson et. al."
date: 
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
  ### global reproducibility of this output
  set.seed(12345)
```


```{r}
plot.custom <- F  ### leave as False for regular pdf report
if (plot.custom) {
  setwd("C:/Users/Paul Gustafson/ownCloud/RES_Dilution")
  source("C:/Users/Paul Gustafson/ownCloud/OPUS2/Rcode/AuxFunctions.R")  
}
```


## Set hyperparameters

```{r}
rng.CFR <- c(0.02, 0.30)

rng.Sn.LC.Carman <- c(0.22, 0.26)

rng.Sn.LC.Potter <- c(0.19, 0.3)

rng.Sn.ILI.Potter <- rng.Sn.ILI.Lemaitre <- c(0.54, 0.78)

rng.Sn.ILI.Hayward <-rng.Sn.ILI.Potter + 0.2 

rng.Sp.LC <- c(1,1)

rng.Sp.ILI.Potter <- rng.Sp.ILI.Lemaitre <- c(0.50, 0.71)

rng.Sp.ILI.Hayward <- c(.3,.71)

rho.ILI <- (-0.95)

rng.a <- c(.07,.3)

rng.k.Carman <- c(1.1, 1.4*1.5)
rng.k.Potter <- c(1.1, 1.6*1.5)
rng.k.Hayward <- c(1.1, 1.2*1.5)
rng.k.Lemaitre <- c(1.1, 1.3*1.5)

### if the study period were 1 year
rng.b.YR <- c(0.11,.17)   

rng.b.Carman <- round((4.5/12)*rng.b.YR, 3)
rng.b.Potter <- round((5/12)*rng.b.YR, 3)
rng.b.Hayward <- rng.b.Lemaitre <- round((2.5/12)*rng.b.YR,3)
```

```{r, echo=F}
tbl <- data.frame(
  Parameter=
  c("Case Fatality Rate", 
    "Sensitivity","","","",
    "Specificity","","",
    "a",
    "k","","","",
    "b","","",""),
  Instance=c("",
             "(Lab-confirmed, Carman)","(Lab-confirmed, Potter)",
             "(ILI Potter, Lemaitre)","(ILI Hayward)",
             "(Lab-confirmed)","(ILI Potter, Lemaitre)","(ILI Hayward)",
             "",
             "(Carman)","(Potter)","(Lemaitre)","(Hayward)",
             "(Carman)","(Potter)","(Lemaitre)","(Hayward)"),             
  rbind(rng.CFR, 
        rng.Sn.LC.Carman, rng.Sn.LC.Potter,   
        rng.Sn.ILI.Potter, rng.Sn.ILI.Hayward,
        rng.Sp.LC, rng.Sp.ILI.Potter, rng.Sp.ILI.Hayward, 
        rng.a, 
        rng.k.Carman, rng.k.Potter, rng.k.Lemaitre, rng.k.Hayward,
        rng.b.Carman, rng.b.Potter, rng.b.Lemaitre, rng.b.Hayward))
names(tbl)[3] <- "Lower"
names(tbl)[4] <- "Upper"
rownames(tbl) <- NULL
```

## Confirm settings in table format

```{r, echo=F}
tbl
```


```{r, echo=F}
if (plot.custom) {
  ### and export table to Word
  library(rtf)
  doc <- RTF("Tbl_prm_rngs.doc")
  addTable(doc, tbl)
  done(doc)
}  
```

## Specify study characteristics

```{r}
specs.carman <- list(
  n.sites=20,
  n.persite=72,
  n.fstr=27,
  y.cn=c(154,688),
  y.tr=c(102,749),
  fstr.cn=c(18,269),
  fstr.tr=c(14,258))

specs.potter.lc <- list(
  n.sites=12,
  n.persite=90,
  n.fstr=20,
  y.cn=c(98,569),
  y.tr=c(50,490),
  fstr.cn=c(6,107),
  fstr.tr=c(5,118))

specs.potter.ili <- specs.potter.lc
specs.potter.ili$n.fstr=90
specs.potter.ili$fstr.cn=c(42,569)
specs.potter.ili$fstr.tr=c(22,490)

specs.lemaitre <- list(
  n.sites=40,
  n.persite=85,
  n.fstr=85,
  y.cn=c(100,1678),
  y.tr=c(89,1722),
  fstr.cn=c(163,1678),
  fstr.tr=c(116,1722))

specs.hayward <- list(
 n.sites=44,
 n.persite=59,
 n.fstr=59, 
 y.cn=c(203,1371),
 y.tr=c(140,1233),
 fstr.cn=c(300,1371),
 fstr.tr=c(142,1233))
```

```{r, echo=F}
tbl <- data.frame(
  c("Carman", "Potter (LC)", "Potter (ILI)", "Lemaitre", "Hayward"),
  rbind(
  unlist(specs.carman), 
  unlist(specs.potter.lc), unlist(specs.potter.ili),
  unlist(specs.lemaitre),  unlist(specs.hayward)))

names(tbl) <- c("Study", "n.fac", "n.per", "n*", "Y.cn", "n", "Y.tr", "n", "F*.cn", "n", "F*.tr", "n")

rownames(tbl) <- NULL

tbl <- data.frame(
rbind(
  c("Carman", specs.carman$n.sites, 
    specs.carman$y.cn[2], "", specs.carman$y.cn[1],
    specs.carman$y.tr[2], "", specs.carman$y.tr[1]),
  
  c("","",specs.carman$fstr.cn[2], specs.carman$fstr.cn[1],"",
          specs.carman$fstr.tr[2], specs.carman$fstr.tr[1],""),
  
  c("Potter (ILI)", specs.potter.ili$n.sites, 
    specs.potter.ili$y.cn[2], specs.potter.ili$fstr.cn[1], specs.potter.ili$y.cn[1],
    specs.potter.ili$y.tr[2], specs.potter.ili$fstr.tr[1], specs.potter.ili$y.tr[1]),
  
  c("(LTI)","",specs.potter.lc$fstr.cn[2], specs.potter.lc$fstr.cn[1],"",
                 specs.potter.lc$fstr.tr[2], specs.potter.lc$fstr.tr[1],""),
  c("Lemaitre", specs.lemaitre$n.sites, 
    specs.lemaitre$y.cn[2], specs.lemaitre$fstr.cn[1],specs.lemaitre$y.cn[1],
    specs.lemaitre$y.tr[2], specs.lemaitre$fstr.tr[1],specs.lemaitre$y.tr[1]),
  c("Hayward", specs.hayward$n.sites, 
    specs.hayward$y.cn[2], specs.hayward$fstr.cn[1],specs.hayward$y.cn[1],
    specs.hayward$y.tr[2], specs.hayward$fstr.tr[1],specs.hayward$y.tr[1])
  )
)

names(tbl) <- c("Study", "# fac.", 
                "# res.", "# infl.","# deaths",
                "# res.", "# infl.","# deaths")
                
rownames(tbl) <- NULL
```

## Confirm characteristics in table format

```{r,echo=F}
print("control columns to left, treatment columns to right")
tbl
```

```{r, echo=F}
if (plot.custom) {
  ### and export it
  library(rtf)
  doc <- RTF("Tbl_study_chars.doc")
  addTable(doc, tbl)
  done(doc)
}  
```

```{r, echo=F} 
### random generation from a symmetric triangluar
### distribution between a and b
rtri <- function(n,a=0,b=1) {
  a + (b-a)*(runif(n)+runif(n))/2  
}

ptri <- function(x, a=0,b=1) {
  md <- (a+b)/2
  ndx <- (x<=md); ans <- rep(NA, length(x))
  ans[ndx] <- 0.5* ( (x[ndx]-a) / (md-a) )^2
  ans[!ndx] <- 1 - 0.5*( (b-x[!ndx]) / (b-md) )^2
ans
}

qtri <- function(p, a=0,b=1) {
  md <- (a+b)/2
  ndx <- (p<=0.5); ans <- rep(NA, length(p))
  ans[ndx] <- a + sqrt(2*p[ndx])*(md-a)
  ans[!ndx] <- b - sqrt(2*(1-p[!ndx]))*(b-md)
ans
}
```


## Function to generate the Step 1 plot

Input is an ensemble of CFR values

```{r}
FirstPlot <- function(cfr, dat.obs, xyl=c(-.05,.15)) {

  dy.hat <- dat.obs$y.cn[1]/dat.obs$y.cn[2] -
            dat.obs$y.tr[1]/dat.obs$y.tr[2]

  dfstr.hat <- dat.obs$fstr.cn[1]/dat.obs$fstr.cn[2] -
               dat.obs$fstr.tr[1]/dat.obs$fstr.tr[2]            

  plot(-2,-2, xlim=xyl, ylim=xyl, 
       xlab=expression(Delta(F)), 
       ylab=expression(Delta(Y)),
       main="(1)")
  polygon(c(0,1,1), c(0,0,1), col=gray(0.75),border=NA)
  abline(v=0); abline(h=0)
  points(dfstr.hat,dy.hat, pch=22)

  for (slp in quantile(cfr, (1:15)/16)) {
    abline(c(0, slp))
  }
}
```

## Function to generate the Step 2 plot

**Additional** inputs are ensembles of sensitivity and specificity values

```{r}
SecondPlot <- function(cfr, sn, sp, dat.obs, xyl=c(-.05,.15)) {
  dy.hat <- dat.obs$y.cn[1]/dat.obs$y.cn[2] -
            dat.obs$y.tr[1]/dat.obs$y.tr[2]
  dfstr.hat <- dat.obs$fstr.cn[1]/dat.obs$fstr.cn[2] -
               dat.obs$fstr.tr[1]/dat.obs$fstr.tr[2]      
  plot(-2,-2, xlim=xyl, ylim=xyl, 
     xlab=expression(Delta(F["*"])),
     ylab=expression(Delta(Y)), 
     main="(2)")
  polygon(c(0,1,1),c(0,0,1), col=gray(.75),border=NA)
  abline(v=0); abline(h=0)
  points(dfstr.hat,dy.hat, pch=22)

  for (slp in quantile(cfr/(sn+sp-1), (1:15)/16)) {
    abline(c(0, slp))
  }  
}
```

## Function to generate the Step 3 plot

**Additional** inputs are ensembles of $a$ and $k$ values (giving rise to $\Delta_F$ values).

```{r}
ThirdPlot <- function(cfr, sn, sp, deltf,
                  dat.obs, xyl=c(-.05,.15)) {

  dy.hat <- dat.obs$y.cn[1]/dat.obs$y.cn[2] -
            dat.obs$y.tr[1]/dat.obs$y.tr[2]
  dfstr.hat <- dat.obs$fstr.cn[1]/dat.obs$fstr.cn[2] -
               dat.obs$fstr.tr[1]/dat.obs$fstr.tr[2]  
  plot(-2,-2, xlim=xyl, ylim=xyl, 
     xlab=expression(Delta(F["*"])),
     ylab=expression(Delta(Y)),
     main="(3)")
  polygon(c(0,1,1),c(0,0,1), col=gray(.75),border=NA)
  abline(v=0)
  abline(h=0)
  points(dfstr.hat,dy.hat, pch=22)
  
  for (i in 1:length(cfr)) {
    points((sn[i]+sp[i]-1)*deltf[i], cfr[i]*deltf[i], cex=.25, pch=20)
  }  
}
```

## Function to generate the Step 4 plot

**Additional** input is an ensemble of b values

```{r}
FourthPlot <- function(cfr, sn, sp, a, k, b, specs,
                       xyl=c(-0.05,.15)) {

  dy.hat <- specs$y.cn[1]/specs$y.cn[2] -
            specs$y.tr[1]/specs$y.tr[2]
  dfstr.hat <- specs$fstr.cn[1]/specs$fstr.cn[2] -
               specs$fstr.tr[1]/specs$fstr.tr[2] 

  plot(-2,-2, xlim=xyl, ylim=xyl, 
     xlab=expression(paste("Est. ", Delta(F["*"]))), 
     ylab=expression(paste("Est. ", Delta(Y))),
     main="(4)")
  polygon(c(0,1,1),c(0,0,1), col=gray(.75),border=NA)
  abline(v=0); abline(h=0)
  points(dfstr.hat,dy.hat, pch=22)

  for (i in 1:length(cfr)) {
    
    smry.fstr <- smry.f <- smry.y <- rep(0,4) 

    ### control facilities 
    for (j in 1:(specs$n.sites/2)) {
      f <- rbinom(specs$n.persite, size=1, prob=rbeta(1, 1, 1/a[i]-1))
      fstr <- rbinom(specs$n.persite, size=1, prob=(1-f)*(1-sp[i])+f*sn[i])
      y <- rbinom(specs$n.persite, size=1, prob=runif(1, 0, b[i])+f*cfr[i])
      smry.f <- smry.f + c(0,0,specs$n.fstr,sum(f[1:specs$n.fstr]))
      smry.fstr <- smry.fstr + c(0,0,specs$n.fstr,sum(fstr[1:specs$n.fstr]))
      smry.y <- smry.y + c(0,0,specs$n.persite,sum(y))
    }  
  
    ### case facilities 
    for (j in 1:(specs$n.sites/2)) {
      f <- rbinom(specs$n.persite, size=1, prob=rbeta(1, 1, k[i]/a[i]-1))
      fstr <- rbinom(specs$n.persite, size=1, prob=(1-f)*(1-sp[i])+f*sn[i])
      y <- rbinom(specs$n.persite, size=1, prob=runif(1, 0, b[i])+f*cfr[i])
      smry.f <- smry.f + c(specs$n.fstr,sum(f[1:specs$n.fstr]),0,0)
      smry.fstr <- smry.fstr + c(specs$n.fstr,sum(fstr[1:specs$n.fstr]),0,0)
      smry.y <- smry.y + c(specs$n.persite,sum(y),0,0)  
    }

    points(smry.fstr[4]/smry.fstr[3] - smry.fstr[2]/smry.fstr[1],
           smry.y[4]/smry.y[3] - smry.y[2]/smry.y[1],cex=.25, pch=20)
  }
}  
```


## Carry out the Carman assessment

```{r}
### number of Monte Carlo draws
n.rep <- 1500

### axis range for these plots
xyl <- c(-0.1,0.14)

### risk difference in mortality (flu minus no-flu)
cfr <- rtri(n.rep, rng.CFR[1], rng.CFR[2])

### now can do first plot

### sens/spec for flu surrogate
sn <- rtri(n.rep, rng.Sn.LC.Carman[1], rng.Sn.LC.Carman[2])
sp <- rtri(n.rep, rng.Sp.LC[1], rng.Sp.LC[2])

### now can do second plot

### a and k values
a <- rtri(n.rep, rng.a[1],rng.a[2])
k <- rtri(n.rep, rng.k.Carman[1],rng.k.Carman[2])

### now can do third plot

### b values
b <- rtri(n.rep, rng.b.Carman[1], rng.b.Carman[2])

### now can do fourth plot
```

```{r, echo=F}

### do the plotting

if (plot.custom) {
  pdf.PG("Fig_Carman.pdf",2,2, ttlspc=T)
  ### and extra nudge to outer margins
  par(oma=.1+c(0,0,0,0.2))
  ### nudge down the axis numbering
  par(cex.axis=0.75)
} else {
  par(mfrow=c(2,2))
}

FirstPlot(cfr, xyl=xyl, 
  dat.obs=specs.carman[c("y.cn","y.tr","fstr.cn","fstr.tr")])

SecondPlot(cfr, sn, sp, xyl=xyl, 
  dat.obs=specs.carman[c("y.cn","y.tr","fstr.cn","fstr.tr")])

ThirdPlot(cfr, sn, sp, deltf=a*(1-1/k), xyl=xyl,
  dat.obs=specs.carman[c("y.cn","y.tr","fstr.cn","fstr.tr")])

FourthPlot(cfr, sn, sp, a, k, b, xyl=xyl, 
  specs=specs.carman)

graphics.off()
```


## Potter, lab-confirmed influenza

Same sequence of four plots 
as previously, but with the following settings changed.


```{r}
sn <- rtri(n.rep, rng.Sn.LC.Potter[1], rng.Sn.LC.Potter[2])
k <- rtri(n.rep, rng.k.Potter[1], rng.k.Potter[2])
b <- rtri(n.rep, rng.b.Potter[1], rng.b.Potter[2])
xyl <- c(-0.10,0.17) ### plotting range
```


```{r, echo=F}

if (plot.custom) {
  pdf.PG("Fig_Potter_LC.pdf",2,2, ttlspc=T)
  ### and extra nudge to outer margins
  par(oma=.1+c(0,0,0,0.2))
  ### nudge down the axis numbering
  par(cex.axis=0.75)
} else {
  par(mfrow=c(2,2))
}

FirstPlot(cfr, xyl=xyl, 
  dat.obs=specs.potter.lc[c("y.cn","y.tr","fstr.cn","fstr.tr")])

SecondPlot(cfr, sn, sp, xyl=xyl, 
  dat.obs=specs.potter.lc[c("y.cn","y.tr","fstr.cn","fstr.tr")])

ThirdPlot(cfr, sn, sp, deltf=a*(1-1/k), xyl=xyl,
  dat.obs=specs.potter.lc[c("y.cn","y.tr","fstr.cn","fstr.tr")])

FourthPlot(cfr, sn, sp, a, k, b, xyl=xyl,
  specs=specs.potter.lc)

graphics.off()
```

## Potter, ILI

All the settings are as above, except for

```{r}
nrm1 <- rnorm(n.rep); nrm2 <- rho.ILI*nrm1 + sqrt(1-rho.ILI^2)*rnorm(n.rep)
sn <- qtri(pnorm(nrm1), rng.Sn.ILI.Potter[1], rng.Sn.ILI.Potter[2])
sp <- qtri(pnorm(nrm2), rng.Sp.ILI.Potter[1], rng.Sp.ILI.Potter[2])

xyl <- c(-0.12,0.18) ### axis range
```


```{r, echo=F}
if (plot.custom) {
  pdf.PG("Fig_Potter_ILI.pdf",2,2, ttlspc=T)
  ### and extra nudge to outer margins
  par(oma=.1+c(0,0,0,0.2))
  ### nudge down the axis numbering
  par(cex.axis=0.75)
} else {
  par(mfrow=c(2,2))
}

FirstPlot(cfr, xyl=xyl, 
  dat.obs=specs.potter.ili[c("y.cn","y.tr","fstr.cn","fstr.tr")])

SecondPlot(cfr, sn, sp, xyl=xyl, 
  dat.obs=specs.potter.ili[c("y.cn","y.tr","fstr.cn","fstr.tr")])

ThirdPlot(cfr, sn, sp, deltf=a*(1-1/k), xyl=xyl,
  dat.obs=specs.potter.ili[c("y.cn","y.tr","fstr.cn","fstr.tr")])

FourthPlot(cfr, sn, sp, a, k, b, xyl=xyl,
  specs=specs.potter.ili)

graphics.off()
```

## Lemaitre, ILI

All the settings are as above, except for

```{r}
k <- rtri(n.rep, rng.k.Lemaitre[1], rng.k.Lemaitre[2])
b <- rtri(n.rep, rng.b.Lemaitre[1], rng.b.Lemaitre[2])
xyl <- c(-0.07,0.11) ### axis range
```


```{r, echo=F}
if (plot.custom) {
  pdf.PG("Fig_Lemaitre.pdf",2,2, ttlspc=T)
  ### and extra nudge to outer margins
  par(oma=.1+c(0,0,0,0.2))
  ### nudge down the axis numbering
  par(cex.axis=0.75)
} else {
  par(mfrow=c(2,2))
}

FirstPlot(cfr, xyl=xyl, 
  dat.obs=specs.lemaitre[c("y.cn","y.tr","fstr.cn","fstr.tr")])

SecondPlot(cfr, sn, sp, xyl=xyl, 
  dat.obs=specs.lemaitre[c("y.cn","y.tr","fstr.cn","fstr.tr")])

ThirdPlot(cfr, sn, sp, deltf=a*(1-1/k), xyl=xyl,
  dat.obs=specs.lemaitre[c("y.cn","y.tr","fstr.cn","fstr.tr")])

FourthPlot(cfr, sn, sp, a, k, b, xyl=xyl,
  specs=specs.lemaitre)

graphics.off()
```

## Hayward, ILI

All the settings are as above, except for

```{r}
nrm1 <- rnorm(n.rep); nrm2 <- rho.ILI*nrm1 + sqrt(1-rho.ILI^2)*rnorm(n.rep)
sn <- qtri(pnorm(nrm1), rng.Sn.ILI.Hayward[1], rng.Sn.ILI.Hayward[2])
sp <- qtri(pnorm(nrm2), rng.Sp.ILI.Hayward[1], rng.Sp.ILI.Hayward[2])

k <- rtri(n.rep, rng.k.Hayward[1],rng.k.Hayward[2])
b <- rtri(n.rep, rng.b.Hayward[1], rng.b.Hayward[2])
xyl <- c(-0.09,0.15) ### axis range
```

```{r, echo=F}
if (plot.custom) {
  pdf.PG("Fig_Hayward.pdf",2,2, ttlspc=T)
  ### and extra nudge to outer margins
  par(oma=.1+c(0,0,0,0.2))
  ### nudge down the axis numbering
  par(cex.axis=0.75)
} else {
  par(mfrow=c(2,2))
}

FirstPlot(cfr, xyl=xyl,
  dat.obs=specs.hayward[c("y.cn","y.tr","fstr.cn","fstr.tr")])

SecondPlot(cfr, sn, sp, xyl=xyl,  
  dat.obs=specs.hayward[c("y.cn","y.tr","fstr.cn","fstr.tr")])

ThirdPlot(cfr, sn, sp, deltf=a*(1-1/k), xyl=xyl,
  dat.obs=specs.hayward[c("y.cn","y.tr","fstr.cn","fstr.tr")])

FourthPlot(cfr, sn, sp, a, k, b, xyl=xyl,
  specs=specs.hayward)

graphics.off()
```


